---
title: "Marketing Spend Analysis - Does More Marekting Spend equal More Purchases?"
author: "Aditya Sharma"
format:
  html: default
  pdf: default
execute: 
  echo: false
  warning: false
  message: false
editor: visual
---  

```{r}
#Installing all the necessary libraries for running the operation
#| label: setup
#| include: false
library(tidyverse)
library(ggplot2)
library(readr)
library(knitr)
```


```{r}
# Reading the dataset into Positron
data <- read_csv("data/marketing_data.csv")
```


```{r}
# Cleaning the raw data and converting text columns "marketing_spend" and "purchase_amount" to real numbers using as.numeric() for further calculations.
cleaned_data <- data |>
  mutate(
    marketingspend = as.numeric(marketing_spend),
    purchaseamount = as.numeric(purchase_amount)
  ) |>
  filter(!is.na(marketingspend), !is.na(purchaseamount))
```

## INTRODUCTION
This dataset shows how marketing spend affects customer purchases across different segments. The research tests if higher marketing spend leads to higher purchase amounts. By understanding this, we can make better marketing decisions which would then lead to boosted sales. For this analysis, I have used a clean customer campaign data.

## MY RESEARCH QUESTION
Does higher marketing spend per customer lead to higher purchase amounts? My analysis compares the relationship using data visualizations. My goal here is to see if the campaign was effective.

## DATA INTRODUCTION
The dataset contains customer ID, signup date, marketing spend, visits, purchase amount, segment, and notes for each customer. This data shows how marketing investments impact purchases. Cleaning converted spend and purchase columns to numbers and removed missing values. No duplicate customers were found.

## DATASET DESCRIPTION
This dataset contains `r nrow(cleaned_data)` observations (rows) and `r ncol(cleaned_data)` variables (columns). Each row represents one customer from the marketing campaign.

The dataset has 7 variables total. The first 3 columns (`customerid`, `marketingspend`, `purchaseamount`) contain numbers that we use for our main analysis. The other 4 columns (`signupdate`, `visits`, `segment`, `notes`) contain text information about each customer.

```{r}
str(cleaned_data)
```

```{r}
summary_table <- data.frame(
  Variable = c("Marketing Spend", "Purchase Amount"),
  Mean = c(
    mean(cleaned_data$marketingspend, na.rm = TRUE),
    mean(cleaned_data$purchaseamount, na.rm = TRUE)
  ),
  Median = c(
    median(cleaned_data$marketingspend, na.rm = TRUE),
    median(cleaned_data$purchaseamount, na.rm = TRUE)
  )
)
knitr::kable(summary_table)
```

**Table 1** shows the mean and median values for marketing spend and purchase amount.

The mean is higher than median, which means a few customers had very high spend/purchase values that pulled the averages up, causing a skew in the data distribution.

```{r}
ggplot(cleaned_data, aes(x = segment, fill = segment)) +
  geom_bar() +
  labs(
    title = "Customer Segments",
    x = "Segment",
    y = "Count"
  ) +
  scale_fill_manual(values = c("Consumer" = "#2E86AB", "SMB" = "#A23B72", "Enterprise" = "#F18F01")) +
  theme_minimal() +
  theme(legend.position = "none")
```

**Figure 1** shows customer distribution by segment.

Here we can clearly see that most of our customers are consumers, followed my Small and Medium Businesses (SMB) who are then followed by Enterprises. This tells us our marketing campaign targeted mostly individual consumers rather than businesses.

```{r}
ggplot(cleaned_data, aes(x = marketingspend, y = purchaseamount)) +
  geom_point(color = "blue", alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 1.2) +
  labs(
    title = "Marketing Spend vs Purchase Amount",
    x = "Marketing Spend",
    y = "Purchase Amount"
  
  )
```

**Figure 2** shows the relationship between marketing spend and purchase amount.

We clearly see an increasing trend with the help of the trend line,  which means higher marketing spend leads to higher customer purchases (positive correlation)

## RESULTS

**Table 1** shows companies spent **$142** per customer on marketing (median $140), while customers bought **$143** worth of products on average (median $139). The revenue **nearly equals** marketing costs, suggesting the campaign **broke even** (before other expenses).

**Figure 1** shows **most customers are Consumers** (individual buyers, not businesses).

**Figure 2** shows an **upward trend line**. **More marketing spend = higher customer purchases** (positive correlation).
